---
title: "Assignment 1"
author: "Steve Kerr (211924)"
date: "2023-09-25"
output: 
  html_document:
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Parsing XML text data

In this assignment we will access and work with German Parliamentary data, which is available in XML format [here](https://www.bundestag.de/services/opendata) (scroll down) for the last two parliamentary periods. Remember XML format is very like HTML format, and we can parse it using a scraper and CSS selectors. Speeches are contained in `<rede>` elements, which each contain a paragraph element describing the speaker, and paragraph elements recording what they said.

### 1.1 

Choose one of the sessions, and retrieve it using R or Python. 

```{r, warning=FALSE, message=FALSE}
# load packages
pacman::p_load(
  dplyr,
  janitor,
  kableExtra,
  purrr,
  quanteda,
  rvest,
  stringr,
  tibble,
  tidyr,
  tidyselect,
  tidytext,
  xml2
)

# retrieve session from 22 Sep 2023
url <- "https://www.bundestag.de/resource/blob/967706/381c8be4289b1cff2ca33f8d641d2333/20123-data.xml"

# read in xml
xml <- read_xml(url)
```

### 1.2 

Using a scraper, get a list of all the <rede> elements. 

```{r}
# extract <rede> elements
rede_list <- xml |> 
  html_elements("rede")

rede_list
```

### 1.3 

For each element, get the name of the speaker, and a single string containing everything that they said. Put this into a dataframe. 

```{r}
# initialize empty df
rede_df <- tibble(
  speaker = character(),
  text = character()
  )

# for each speech in the list of speeches
for (i in 1:length(rede_list)) {
  
  # extract speaker name
  speaker <- rede_list[i] |> 
    html_elements("titel, vorname, nachname") |> 
    xml_text() |> 
    str_c(collapse = " ")
  
  # extract text
  text <- rede_list[i] |> 
    # include class 'J', 'O', and the first instance of 'J_1' since these are relevant to the speech
    xml_find_all(".//p[@klasse='O'] | .//p[@klasse='J'] | .//p[@klasse='J_1'][position() < last()]") |>
    html_text2() |> 
    str_c(collapse = "\n")
  
  # combine into dataframe
  rede_df <- rede_df |> 
    add_row(speaker = speaker, text = text)
  
}

# print the top three rows of the data frame
rede_df |> 
  head(3) |> 
  kbl() |> 
  column_spec(2, width = "300in") |>  # set max width
  kable_styling(full_width = FALSE) 
```

### 2.1 

Choose a politician, and print the number of speeches they made in this session

```{r}
# I've heard folks mention Steffi Lemke, so let's take a look at her speeches

# count the number of speeches by Frau Lemke
rede_df |> 
  filter(str_detect(speaker, "Steffi Lemke")) |> 
  count(speaker)
```

### 2.2 

Print the content of the first speech by the politician you choose.

```{r}
# print the first speech by Steffi Lemke
rede_df |> 
  filter(str_detect(speaker, "Steffi Lemke")) |> 
  head(1) |> 
  pull(text) |> 
  cat()
```

### 2.3 

Process the list of speeches into a TFIDF matrix. What are the highest scoring terms in this matrix for the first speech by the politician you have chosen?

```{r}
dfm_tidy <- rede_df |> 
  corpus() |> # speech dataframe to corpus
  tokens() |> # to tokens
  tokens_remove(pattern = stopwords("de")) |> # remove German stopwords
  tokens_wordstem(language = "german") |> # stemming
  dfm() |> # to dfm
  dfm_tfidf() |> # calculate TFIDF
  tidy()

# print top 10 words (stems) by TFIDF score
dfm_tidy |> 
  filter(document == paste0("text", which(rede_df$speaker == "Steffi Lemke"))) |> # filter for Frau Lemke's speech
  select(-document) |> 
  arrange(desc(count)) |> 
  head(10) |> 
  kbl() |> 
  kable_styling()
```

### 2.4 

Using the resource "Stammdaten aller Abgeordneten seit 1949 im XML-Format", retrieve the records pertaining to your chosen politician and print the information they contain.

```{r}
# download zip file from url
download.file("https://www.bundestag.de/resource/blob/472878/a4859899e44a7cab1a8233e5dd69f2f3/MdB-Stammdaten-data.zip", "temp")

# unzip file
unzip(here::here("KERR/temp"))

# extract Lemke's record
mdb_raw <- here::here("KERR/MDB_STAMMDATEN.XML") |>
  read_xml() |> 
  xml_find_all("//MDB[NAMEN/NAME/NACHNAME/text() = 'Lemke']") 

# convert to df 
mdb_df <- mdb_raw |> 
  as_list() |> 
  as_tibble_col() |> 
  unnest_wider("value")

# write function to unlist elements of a df (found via Google @jolyphil)
unlist_all <- function(x) { 
  x[sapply(x, is.null)] <- NA
  unlist(x)
}

# extract name info
name_df <- mdb_df |> 
  select(ID, NAMEN) |> 
  unnest_longer("NAMEN", indices_include = FALSE) |> 
  unnest_wider("NAMEN") |> 
  mutate(across(.cols = everything(), .fns = unlist_all)) |> # unlist everything
  clean_names() |> # tidy names
  select_if(~!all(is.na(.))) # drop empty columns

# print name info
name_df |> 
  kbl() |> 
  kable_styling()

# extract bio info
bio_df <- mdb_df |> 
  select(ID, BIOGRAFISCHE_ANGABEN) |> 
  unnest_wider("BIOGRAFISCHE_ANGABEN") |> 
  mutate(across(.cols = everything(), .fns = unlist_all)) |> # unlist everything
  clean_names() |> # tidy names
  select_if(~!all(is.na(.))) # drop empty columns

# print bio info
bio_df |> 
  kbl() |> 
  kable_styling()

# extract election info
elec_df <- mdb_df |>
    select("ID", "WAHLPERIODEN") |>
    unnest_longer("WAHLPERIODEN", indices_include = FALSE) |>
    unnest_wider("WAHLPERIODEN") |>
    mutate(across(.cols = c("ID",
                            "WP",
                            "MDBWP_VON",
                            "MDBWP_BIS",
                            "WKR_NUMMER",
                            "WKR_NAME",
                            "WKR_LAND",
                            "LISTE",
                            "MANDATSART"),
                  .fns = unlist_all)) |> 
    clean_names() |> # tidy names
    select_if(~!all(is.na(.))) # drop empty columns

# print election info
elec_df |> 
  kbl() |> 
  kable_styling()
```
